Modelling BCG vaccination in the UK

### 2019-07-12 rmds/03-getTBinR.Rmd
first pass: getTBinR comments

`getTBinR` is an R package [@RCoreTeam2019] to facilitate working with the data [@WHO:2018] collected by the World Health Organization (WHO) on the country level epidemiology of Tuberculosis (TB). All data is freely available from the [WHO](https://www.who.int/tb/country/data/download/en/) and the package code is archived on Zenodo [@Abbott:2019] and [Github](https://www.samabbott.co.uk/getTBinR/). The aim of `getTBinR` is to allow researchers, and other interested individuals, to quickly and easily gain access to a detailed TB data set and to start using it to derive key insights. It provides a consistent set of tools that can be used to rapidly evaluate hypotheses on a widely used data set before they are explored further using more complex methods or more detailed data. The functions provided in this package were developed to have sensible defaults to allow those new to the field to quickly gain key insights but also allow sufficient customisation so that experienced users may rapidly prototype new ideas.

The `getTBinR` package facilitates downloading the most up-to-date version of multiple TB relevant data sources from the WHO, along with the accompanying data dictionaries. It also contains functions to allow easy exploration of the data via searching data dictionaries, summarising key metrics on a regional and global level, and visualising the data in a variety of highly customisable ways. In addition, it provides both a dashboard and an automated country level report that enables the global, regional, and country level picture to be quickly summarised. An example of a potential use of the package is to explore estimates of the TB case fatality ratio [@Abbott:2018]. In a few lines of code, using only built in package tooling, large regional differences can be discovered. Further insights can then be found by linking to other publicly available data sets or using a model based analysis. See https://www.samabbott.co.uk/getTBinR/ for documentation. The remainder of this chapter outlines the key functionality of `getTBinR 0.6.1` and summarises the current impact.

@@

`getTBinR` has been developed using R package best practices and as such is thoroughly tested  using an automated testing suite that runs against Linux, Windows and MacOS environments. Package documentation is supplied in a searchable website (https://www.samabbott.co.uk/getTBinR/) and a development environment can be launched with a single button press (https://mybinder.org/v2/gh/seabbs/getTBinR/master?urlpath=rstudio). Use cases for the package have been outlined using multiple case studies, see the package documentation for details.

@@

 * As of the 1st of July 2019, `getTBinR` has been released on CRAN for over a year. It has been downloaded over 9000 times, has no outstanding bug-related issues, and has received multiple updates greatly expanding the functionality available. The standalone dashboard hosted online (http://www.seabbs.co.uk/shiny/ExploreGlobalTB/) has had over 3000 unique users.

>>>>

`getTBinR` is an R package[@RCoreTeam2019] that I developed to facilitate working with the data[@WHO:2018] collected by the World Health Organization (WHO) on the country level epidemiology of Tuberculosis (TB). All data is freely available from the WHO^[WHO TB data: https://www.who.int/tb/country/data/download/en/] and the package code is archived on Zenodo[@Abbott:2019] and Github^[GitHub: https://github.com/seabbs/getTBinR/). The aim of `getTBinR` is to allow researchers, and other interested individuals, to quickly and easily gain access to a detailed TB data set and to start using it to derive key insights. It provides a consistent set of tools that can be used to rapidly evaluate hypotheses on a widely used data set before they are explored further using more complex methods or more detailed data. The remainder of this chapter outlines the key functionality of `getTBinR 0.6.1` , and summarises the current impact.

@@

`getTBinR` has been developed using R package best practices and as such is thoroughly tested  using an automated testing suite that runs against Linux, Windows and MacOS environments. Package documentation is supplied in a searchable website^[Website: https://www.samabbott.co.uk/getTBinR/] and a development environment can be launched with a single button press^[Binder: https://mybinder.org/v2/gh/seabbs/getTBinR/master?urlpath=rstudio]. Use cases for the package have been outlined using multiple case studies, see the package documentation for details.

## Discussion

In this chapter I have introduced the `getTBinR` R package. `getTBinR` facilitates downloading the most up-to-date version of multiple TB relevant data sources from the WHO, along with the accompanying data dictionaries. It also contains functions to allow easy exploration of the data via searching data dictionaries, summarising key metrics on a regional and global level, and visualising the data in a variety of highly customisable ways. In addition, it provides both a dashboard and an automated country level report that enables the global, regional, and country level picture to be quickly summarised. It was developed using R package development best practises and has been peer-reviewed.[@Abbott:2018]

As of the 1st of August 2019, `getTBinR` has been released on CRAN for over a year. It has been downloaded over 10,000 times, has a growing user base, no outstanding bug-related issues, and has received multiple updates greatly expanding the functionality available. The standalone dashboard hosted online^[Dashboard: http://www.seabbs.co.uk/shiny/ExploreGlobalTB/] has had over 3000 unique users. It has been used as a teaching aid, as an example of open science, to facilate exploratory data analysis and to provide context for other research. In this thesis it has been used extensively in Chapter \@ref(background) to provide context and was also used as a hypothesis generating tool in all other chapters. Outside of the work presented in this thesis I have used `getTBinR` extensively as a data analysis tool^[Blog: https://www.samabbott.co.uk/tags/who/], to widen the appreciation of TB as a global health problem^[Twitter:https://twitter.com/search?q=getTBinR&src=typd]^[Reddit: https://www.reddit.com/user/seabbs/posts/], and to provide contextual information for other research^[Presentation: https://www.samabbott.co.uk/what-do-we-really-know-about-bcg/presentation.html#1].

Whilst `getTBinR` is feature complete, and stable, development work continues. Future projects include: using the `shinymeta` R package to provide downloadable R code to users of the interactive application; iterating on the current automated report to improve formating and to increase the amount of information displayed. As additional WHO TB data are released they will be added to `getTBinR`.

@@

 * The current impact and future direction of `getTBinR` has also be detailed.
